---
title: "Homework 4"
author: "PSTAT 131/231"
output:
    html_document:
      toc: true
      toc_float: true
      code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE,
                      warning = FALSE)
library(tidymodels)
library(tidyverse)
library(readr)
library(ISLR) 
library(ISLR2) 
library(discrim)
library(poissonreg)
library(corrr)
library(klaR) 
```

### Question 1

Split the data, stratifying on the outcome variable, `survived.`  You should choose the proportions to split the data into. Verify that the training and testing data sets have the appropriate number of observations. 
```{r}
set.seed(3435)
titanic <- read.csv("C:\\Users\\Jonat\\OneDrive\\schoolwork\\PSTAT 131\\HW\\HW 4\\homework-4\\data\\titanic.csv")

titanic$pclass <- as.factor(titanic$pclass)
titanic$survived <- as.factor(titanic$survived)
titanic$survived <- relevel(titanic$survived, 'Yes')
head(titanic)

t_split <- initial_split(titanic, prop = 0.7, strata = survived)
t_train <- training(t_split)
t_test <- testing(t_split)
dim(titanic)
dim(t_train)
dim(t_test)
```

### Question 2

Fold the **training** data. Use *k*-fold cross-validation, with $k = 10$.
```{r}
t_fold <- vfold_cv(t_train, v = 10)
t_fold
```

### Question 3

In your own words, explain what we are doing in Question 2. What is *k*-fold cross-validation? Why should we use it, rather than simply fitting and testing models on the entire training set? If we **did** use the entire training set, what resampling method would that be?

-K-Fold Cross-Validation is the process of trying to find the best possible value of degree that produces the closest polynomial degree value that fits within the regression model. Our models may not be distributed at degree 1. It may have a degree higher than 1 and in that case it is curved. That is why k-fold-cross-validation may be useful. If we use the training set, the resampling method would be the validation set.

### Question 4

Set up workflows for 3 models:

1. A logistic regression with the `glm` engine;
2. A linear discriminant analysis with the `MASS` engine;
3. A quadratic discriminant analysis with the `MASS` engine.

How many models, total, across all folds, will you be fitting to the data? To answer, think about how many folds there are, and how many models you'll fit to each fold.

```{r}
t_recipe <- recipe(survived ~ pclass + sex + age + sib_sp + parch + fare, t_train) %>%
  step_impute_linear(age, impute_with = imp_vars(sib_sp)) %>%
  step_dummy(all_nominal_predictors())

log_reg <- logistic_reg() %>%
  set_engine("glm") %>%
  set_mode("classification")
log_wkflow <- workflow() %>%
  add_model(log_reg) %>%
  add_recipe(t_recipe)

lda_mod <- discrim_linear() %>%
  set_engine("MASS") %>%
  set_mode("classification")
lda_wkflow <- workflow() %>%
  add_model(lda_mod) %>%
  add_recipe(t_recipe)

qda_mod <- discrim_quad() %>%
  set_engine("MASS") %>%
  set_mode("classification")
qda_wkflow <- workflow() %>%
  add_model(qda_mod) %>%
  add_recipe(t_recipe)
```

### Question 5

Fit each of the models created in Question 4 to the folded data.

**IMPORTANT:** *Some models may take a while to run – anywhere from 3 to 10 minutes. You should NOT re-run these models each time you knit. Instead, run them once, using an R script, and store your results; look into the use of [loading and saving](https://www.r-bloggers.com/2017/04/load-save-and-rda-files/). You should still include the code to run them when you knit, but set `eval = FALSE` in the code chunks.*

```{r}
log_folded <- log_wkflow %>%
  fit_resamples(t_fold)
lda_folded <- lda_wkflow %>%
  fit_resamples(t_fold)
qda_folded <- qda_wkflow %>%
  fit_resamples(t_fold)
```

### Question 6

Use `collect_metrics()` to print the mean and standard errors of the performance metric *accuracy* across all folds for each of the four models.

Decide which of the 3 fitted models has performed the best. Explain why. *(Note: You should consider both the mean accuracy and its standard error.)*

```{r}
collect_metrics(log_folded)
collect_metrics(lda_folded)
collect_metrics(qda_folded)
```
Log had the highest accuracy and the lowest standard error.

### Question 7

Now that you’ve chosen a model, fit your chosen model to the entire training dataset (not to the folds).
```{r}
log_fit <- fit(log_wkflow, t_train)
```
### Question 8

Finally, with your fitted model, use `predict()`, `bind_cols()`, and `accuracy()` to assess your model’s performance on the testing data!

Compare your model’s testing accuracy to its average accuracy across folds. Describe what you see.
```{r}
t_train_log <- predict(log_fit, new_data = t_train, type = 'prob')
t_train_log <- bind_cols(t_train_log)
log_acc <- augment(log_fit, new_data = t_train) %>%
  accuracy(truth = survived, estimate = .pred_class)
log_acc
```
The model accuracy on the training data set has an accuracy of about 0.7913323. This is close to the accuracy of the three folded models. All the models were around the 0.79 mark. This means that our folded models performed decently well in fitting the training model.

## Required for 231 Students

Consider the following intercept-only model, with $\epsilon \sim N(0, \sigma^2)$:

$$
Y=\beta+\epsilon
$$

where $\beta$ is the parameter that we want to estimate. Suppose that we have $n$ observations of the response, i.e. $y_{1}, ..., y_{n}$, with uncorrelated errors.

### Question 9

Derive the least-squares estimate of $\beta$.

### Question 10

Suppose that we perform leave-one-out cross-validation (LOOCV). Recall that, in LOOCV, we divide the data into $n$ folds. What is the covariance between $\hat{\beta}^{(1)}$, or the least-squares estimator of $\beta$ that we obtain by taking the first fold as a training set, and $\hat{\beta}^{(2)}$, the least-squares estimator of $\beta$ that we obtain by taking the second fold as a training set?